\specchapter{Program environment}

\specitem
The implementation translates source files and executes programs in two phases,
respectively referred to as the \textit{translation phase} and the
\textit{execution phase}. The context in which these phases occur is referred
to as the \textit{translation environment} and the \textit{execution
environment}.

\specsection{Translation environment}

\specsubitem
A Hare program consists of one or more \textit{source files} which are provided
to the translation phase. A source file shall be represented as \hbox{UTF-8}
text.

\specsubitem
Each \textit{source file} is a member of exactly one \textit{module}, and the
collective source files for a module form a \textit{translation unit}. Each
module may define its own private types, data, procedures, and so on, for the
purpose of accomplishing its tasks. It may also \textit{export} these types for
other modules to use.

\specsubitem
A Hare program may be translated incrementally, rather than all at one time.
However, the composite source files of a single translation unit must be
translated together. Once translated, an implementation-defined representation
of their \textit{exported interface} (the identifiers and types of their
exported declarations) provides sufficient information to translate other units
which require the module. The translated modules may be composed into the final
program at the last step.

\specsubitem
If the source files for a translation unit are not changed, the translated
module may be used repeatedly without repeating the translation step.

\specsection{Translation steps}

\specsubitem
The list of source files constituting the translation unit are identified. Steps
\subitemref{Translation steps}{2} and \subitemref{Translation steps}{3} are
repeated for each source file.

\specsubitem
Lexical analysis is conducted on the source file, translating it into a stream
of \textit{tokens}.

Forward references: \secref{Lexical analysis}

\specsubitem
Syntax analysis is conducted on the token stream, mapping the tokens to a
\nonterminal{sub-unit}.

Forward references: \secref{Units}

\specsubitem
Logical analysis is conducted on the sub-units. In this step, the implementation
verifies the constraints imposed on the program. The result of this step is a
\textit{verified program module}.

\informative{In this step, colloquially referred to as the "check" step, a
module composed of several source files is consolidated into a single verified
program module.}

\specsubitem
Once the verified program module is obtained in the translation phase, the
remainder of the translation phase shall be completed with no further
diagnostic messages, except in the case where external factors from the
execution environment prevent successful completion.

\example{Memory exhaustion or lack of disk space are situations which may cause
a failure in the remainder of the translation process.}

\specsubitem
The verified program module is combined with any applicable external modules and
translated into a single program image which is suitable for interpretation by
the execution environment.

\specsection{Execution environment}

\specsubitem
Three execution environments are defined: \textit{freestanding},
\textit{hosted}, and \textit{test}. The implementation must support a
freestanding environment; support for any other environment is optional.

\specsubitem
During \textit{program startup}, the execution environment shall first
initialize all global declarations to their initial values in an
implementation-defined manner. The next steps are dependent on the execution
environment.

\specsubitem
Behavior during \textit{program teardown} is dependent on the execution
environment.

\specsubsection{The freestanding environment}

\specsubsubitem
In the freestanding environment, behavior during program startup after global
declarations are initialized, behavior during program teardown, as well as
behavior of initialization and finalization functions, is undefined.

Forward references: \subsecref{Function declarations}

\specsubsection{The hosted environment}

\specsubsubitem
In the hosted environment, after global declarations are initialized during
program startup, the execution environment shall call all initialization
functions in an order such that the initialization functions for a given
module are run before those of any module that depends on it, then transfer
control to the program \textit{entry point} in an implementation-defined manner.
The ordering of initialization functions within a module shall be undefined.

\specsubsubitem
Program teardown in the hosted environment shall cause the execution environment
to call all finalization functions in an order such that the finalization
functions for a given module are run after those of any module that depends on
it, then terminate. The ordering of finalization functions within a module shall
be undefined.

\specsubsubitem
The program entry point shall be a function named \code{main} in the root
namespace. The declaration shall not have any attributes. The function shall
have no parameters and a result type of \terminal{void}. The declaration shall
be exported, unless the declaration is a prototype, in which case it need not be
exported.

\informative{The signature of a conformant entry point follows:}

\begin{codesample}
export fn main() void;
\end{codesample}

\informative{The program shall provide this declaration in the root namespace.}

\specsubsubitem
Alternatively, the entry point may be declared with a different name and/or in a
different namespace if a recognized form of \terminal{@symbol} is supplied to
the declaration. The set of recognized forms, if any exist, are
implementation-defined.

\informative{If an implementation used the symbol "main" for its entry point,
the following signature would be a valid entry point declaration:}

\begin{codesample}
export @symbol("main") fn any::other::name() void;
\end{codesample}

\specsubsubitem
The identifier \code{main} shall not be used for any root namespace declaration
which would not be a valid entry point, unless the declaration uses any of
\terminal{@symbol}, \terminal{@init}, \terminal{@fini}, or \terminal{@test}.
The implementation-defined set of recognized forms of \terminal{@symbol} shall
not be used in any namespace for a declaration which would not be a valid entry
point.

\specsubsubitem
Returning from the program entry point shall begin the program teardown process.

Forward references: \subsecref{Global declarations}, \subsecref{Function declarations}

\specsubsection{The test environment}

\specsubsubitem
In the test environment, after global declarations are initialized during
program startup, the execution environment shall call all initialization
functions in an order such that the initialization functions for a given
module are run before those of any module that depends on it, then transfer
control to the \textit{test runner}. The ordering of initialization functions
within a module shall be undefined.

\specsubsubitem
Program teardown in the hosted environment shall cause the execution environment
to call all finalization functions in an order such that the finalization
functions for a given module are run after those of any module that depends on
it, then terminate. The ordering of initialization functions within a module
shall be undefined.

\specsubsubitem
The test runner shall call any number of test functions in an unspecified order,
and then begin the program teardown process.

\informative{The implementation should provide a means for the user to specify
which tests to run.}

\specsubsubitem
Prior to calling each test function, as well as after calling the final test
function, the test runner shall reset the execution environment in an
implementation-defined manner.

\example{Resetting the execution environment may involve setting the floating
point rounding mode to the default, clearing all floating point exceptions,
resetting all signal handlers to their defaults, and routing standard I/O
streams to known locations.}

\specsubsubitem
Each test function designates one test, which may either \textit{pass} or
\textit{fail}. By default, if control returns from a test function, the test
passes; if the test function aborts (such as from an
\nonterminal{assertion-expression} or a failed type assertion), the test fails.
The implementation may declare functions in the runtime library which change the
pass and fail conditions for the running test. Aborting from a test function
shall not terminate the program; the next test function shall always be
executed.

\specsubsubitem
The test runner shall keep track of which tests pass and which tests fail, and
print this information in an implementation-defined manner, possibly alongside
other messages.

Forward references: \subsecref{Function declarations}

\specsubsection{Program execution}

\specsubsubitem
The evaluation of an expression may have \textit{side-effects} in addition to
computing a value. Calling a function or modifying an object is considered a
side-effect.

\specsubsubitem
If the implementation is able to determine that the evaluation of part of an
expression is not necessary to compute the correct value and cause the same
side-effects to occur in the same order, it may rewrite or re-order the
expressions or sub-expressions to produce the same results more optimally.

\informative{The interpretation of this constraint should be conservative.
Implementations should prefer to be predictable over being fast. Programs which
require greater performance should prefer to hand-optimize their source code for
this purpose.}

Forward references: \secref{Expressions}

\specsection{Diagnostics}

\specsubitem
If the constraints are found to be invalid during the translation phase, the
implementation shall display an error indicating which constraint was
invalidated, and indicate that the translation has failed in whatever manner is
semantically appropriate.

\example{On a Unix-like system, the semantically appropriate indication of
failure is to exit with a non-zero status code.}

\specsubitem
In the translation environment, if the implementation is able to determine that
multiple constraints are invalid, it may display several diagnostic messages.

\specsubitem
If the constraints are found to be invalid during the execution phase, a hosted
implementation shall abort the execution phase, display a diagnostic message,
and indicate that the execution has failed in whatever manner is semantically
appropriate.
